---
title: An R Markdown document converted from "HW3_CIA_World_Factbook.ipynb"
output: html_document
---

# Homework 3: CIA World Factbook Data Cleaning and Formatting

**Author:** Amogh Guthur
**Course:** DACSS 601

## Data Sources:
- Maternal Mortality Ratio: https://www.cia.gov/the-world-factbook/field/maternal-mortality-ratio/country-comparison/
- Infant Mortality Rate: https://www.cia.gov/the-world-factbook/field/infant-mortality-rate/country-comparison/
- Life Expectancy at Birth: https://www.cia.gov/the-world-factbook/field/life-expectancy-at-birth/country-comparison/

## Task:
Keep only 3 columns: **country name**, **variable**, and **region**. Save as RDS and CSV.

## 1. Maternal Mortality Ratio Data

```{r}
# Create maternal mortality ratio dataframe with 3 columns: country, variable, region
maternal_mortality <- data.frame(
  # Column 1: Country names from CIA World Factbook
  country = c("Nigeria", "Chad", "South Sudan", "Central African Republic", "Liberia",
              "Somalia", "Afghanistan", "Benin", "Guinea-Bissau", "Guinea",
              "Lesotho", "Madagascar", "Congo, Democratic Republic of the", "Burundi", "Mauritania",
              "Kenya", "Mali", "Cote d'Ivoire", "Zimbabwe", "Gambia, The",
              "Sierra Leone", "Niger", "Togo", "Haiti", "Eritrea",
              "Tanzania", "Nauru", "Cameroon", "Sudan", "Burkina Faso",
              "Congo, Republic of the", "Senegal", "Ghana", "Gabon", "Rwanda",
              "Venezuela", "Malawi", "Ethiopia", "Timor-Leste", "Papua New Guinea",
              "Burma", "Angola", "Comoros", "Equatorial Guinea", "Uganda",
              "Tuvalu", "Djibouti", "Pakistan", "Marshall Islands", "Botswana",
              "Bolivia", "Nepal", "Indonesia", "Namibia", "Cambodia",
              "Jamaica", "Micronesia, Federated States of", "Dominican Republic", "Solomon Islands", "South Africa",
              "Eswatini", "Yemen", "Bangladesh", "Laos", "Samoa",
              "Vanuatu", "Guatemala", "Palau", "Zambia", "Philippines",
              "Suriname", "Mozambique", "Kiribati", "India", "Bahamas, The",
              "Guyana", "Sao Tome and Principe", "Saint Kitts and Nevis", "Morocco", "Belize",
              "Korea, North", "Tonga", "Brazil", "Iraq", "Mauritius",
              "Algeria", "Nicaragua", "Colombia", "Libya", "Paraguay",
              "Saint Vincent and the Grenadines", "Ecuador", "Trinidad and Tobago", "Peru", "Grenada",
              "Vietnam", "Bhutan", "Honduras", "Saint Lucia", "Kyrgyzstan",
              "Mexico", "Seychelles", "Mongolia", "Cabo Verde", "El Salvador",
              "Panama", "Dominica", "Tunisia", "Brunei", "Antigua and Barbuda",
              "Cuba", "Barbados", "Thailand", "Argentina", "Maldives",
              "Jordan", "Fiji", "Uzbekistan", "Malaysia", "Costa Rica",
              "Georgia", "Syria", "Latvia", "Moldova", "Armenia",
              "Sri Lanka", "Azerbaijan", "Egypt", "Bahrain", "United States",
              "China", "West Bank", "Gaza Strip", "Iran", "Turkey (Turkiye)",
              "Uruguay", "Portugal", "Ukraine", "Lebanon", "Cyprus",
              "Tajikistan", "Oman", "Canada", "Romania", "Luxembourg",
              "Hungary", "Andorra", "Serbia", "Puerto Rico", "Chile",
              "Kazakhstan", "Russia", "Lithuania", "Kuwait", "United Kingdom",
              "Finland", "Malta", "San Marino", "France", "Albania",
              "New Zealand", "Saudi Arabia", "Italy", "Bosnia and Herzegovina", "Montenegro",
              "Singapore", "Austria", "Bulgaria", "Turkmenistan", "Estonia",
              "Greece", "Switzerland", "Monaco", "Ireland", "Qatar",
              "Netherlands", "Belgium", "Denmark", "Korea, South", "Sweden",
              "Slovakia", "Germany", "Japan", "Spain", "North Macedonia",
              "Iceland", "Czechia", "Croatia", "United Arab Emirates", "Slovenia",
              "Australia", "Israel", "Poland", "Belarus", "Norway",
              "Cook Islands"),
  # Column 2: Maternal mortality ratio (deaths per 100,000 live births, 2023 est.)
  maternal_mortality_ratio = c(993, 748, 692, 692, 628,
                               563, 521, 518, 505, 494,
                               478, 445, 427, 392, 381,
                               379, 367, 359, 358, 354,
                               354, 350, 349, 328, 291,
                               276, 273, 258, 256, 242,
                               241, 237, 234, 233, 229,
                               227, 225, 195, 192, 189,
                               185, 183, 179, 174, 170,
                               170, 162, 155, 155, 155,
                               146, 142, 140, 139, 137,
                               130, 129, 124, 123, 118,
                               118, 118, 115, 112, 101,
                               100, 94, 89, 85, 84,
                               84, 82, 80, 80, 76,
                               75, 75, 74, 70, 67,
                               67, 67, 67, 66, 66,
                               62, 60, 59, 59, 58,
                               56, 55, 54, 51, 48,
                               48, 47, 47, 44, 42,
                               42, 42, 41, 40, 39,
                               37, 36, 36, 36, 35,
                               35, 35, 34, 33, 32,
                               31, 30, 26, 26, 24,
                               20, 20, 19, 19, 19,
                               18, 18, 17, 17, 17,
                               16, 16, 16, 16, 15,
                               15, 15, 15, 15, 14,
                               14, 13, 12, 12, 12,
                               12, 11, 11, 11, 10,
                               10, 9, 8, 8, 8,
                               8, 8, 8, 7, 7,
                               7, 7, 6, 6, 6,
                               6, 6, 6, 5, 5,
                               5, 5, 5, 4, 4,
                               4, 4, 4, 4, 4,
                               4, 4, 3, 3, 3,
                               3, 3, 3, 3, 3,
                               2, 2, 2, 1, 1,
                               0),
  # Column 3: Region classification from CIA World Factbook
  region = c("Africa", "Africa", "Africa", "Africa", "Africa",
             "Africa", "South Asia", "Africa", "Africa", "Africa",
             "Africa", "Africa", "Africa", "Africa", "Africa",
             "Africa", "Africa", "Africa", "Africa", "Africa",
             "Africa", "Africa", "Africa", "Central America and the Caribbean", "Africa",
             "Africa", "Australia and Oceania", "Africa", "Africa", "Africa",
             "Africa", "Africa", "Africa", "Africa", "Africa",
             "South America", "Africa", "Africa", "East and Southeast Asia", "Australia and Oceania",
             "East and Southeast Asia", "Africa", "Africa", "Africa", "Africa",
             "Australia and Oceania", "Africa", "South Asia", "Australia and Oceania", "Africa",
             "South America", "South Asia", "East and Southeast Asia", "Africa", "East and Southeast Asia",
             "Central America and the Caribbean", "Australia and Oceania", "Central America and the Caribbean", "Australia and Oceania", "Africa",
             "Africa", "Middle East", "South Asia", "East and Southeast Asia", "Australia and Oceania",
             "Australia and Oceania", "Central America and the Caribbean", "Australia and Oceania", "Africa", "East and Southeast Asia",
             "South America", "Africa", "Australia and Oceania", "South Asia", "Central America and the Caribbean",
             "South America", "Africa", "Central America and the Caribbean", "Africa", "Central America and the Caribbean",
             "East and Southeast Asia", "Australia and Oceania", "South America", "Middle East", "Africa",
             "Africa", "Central America and the Caribbean", "South America", "Africa", "South America",
             "Central America and the Caribbean", "South America", "Central America and the Caribbean", "South America", "Central America and the Caribbean",
             "East and Southeast Asia", "South Asia", "Central America and the Caribbean", "Central America and the Caribbean", "Central Asia",
             "North America", "Africa", "East and Southeast Asia", "Africa", "Central America and the Caribbean",
             "Central America and the Caribbean", "Central America and the Caribbean", "Africa", "East and Southeast Asia", "Central America and the Caribbean",
             "Central America and the Caribbean", "Central America and the Caribbean", "East and Southeast Asia", "South America", "South Asia",
             "Middle East", "Australia and Oceania", "Central Asia", "East and Southeast Asia", "Central America and the Caribbean",
             "Middle East", "Middle East", "Europe", "Europe", "Middle East",
             "South Asia", "Middle East", "Africa", "Middle East", "North America",
             "East and Southeast Asia", "Middle East", "Middle East", "Middle East", "Middle East",
             "South America", "Europe", "Europe", "Middle East", "Europe",
             "Central Asia", "Middle East", "North America", "Europe", "Europe",
             "Europe", "Europe", "Europe", "Central America and the Caribbean", "South America",
             "Central Asia", "Europe", "Europe", "Middle East", "Europe",
             "Europe", "Europe", "Europe", "Europe", "Europe",
             "Australia and Oceania", "Middle East", "Europe", "Europe", "Europe",
             "East and Southeast Asia", "Europe", "Europe", "Central Asia", "Europe",
             "Europe", "Europe", "Europe", "Europe", "Middle East",
             "Europe", "Europe", "Europe", "East and Southeast Asia", "Europe",
             "Europe", "Europe", "East and Southeast Asia", "Europe", "Europe",
             "Europe", "Europe", "Europe", "Middle East", "Europe",
             "Australia and Oceania", "Middle East", "Europe", "Europe", "Europe",
             "Australia and Oceania")
)
```

```{r}
# Display first 6 rows to verify structure
head(maternal_mortality)
```

```{r}
# Verify data types using str() function
str(maternal_mortality)
```

```{r}
# Save as RDS format (preserves R data types)
saveRDS(maternal_mortality, "maternal_mortality.rds")
```

```{r}
# Save as CSV format (plain text, portable)
write.csv(maternal_mortality, "maternal_mortality.csv", row.names = FALSE)
```

```{r}
# Verify RDS file by reading it back and checking structure
mm_rds <- readRDS("maternal_mortality.rds")
# Display structure of loaded RDS file
str(mm_rds)
```

```{r}
# Verify CSV file by reading it back and checking structure
mm_csv <- read.csv("maternal_mortality.csv")
# Display structure of loaded CSV file
str(mm_csv)
```

## 2. Infant Mortality Rate Data

```{r}
# Create infant mortality rate dataframe with 3 columns: country, variable, region
infant_mortality <- data.frame(
  # Column 1: Country names from CIA World Factbook
  country = c("Afghanistan", "Somalia", "Central African Republic", "Equatorial Guinea", "Sierra Leone",
              "Niger", "Chad", "South Sudan", "Mozambique", "Mali",
              "Congo, Democratic Republic of the", "Liberia", "Angola", "Comoros", "Nigeria",
              "Benin", "Cote d'Ivoire", "Pakistan", "Mauritania", "Burkina Faso",
              "Guinea", "Guinea-Bissau", "Cameroon", "Lesotho", "Djibouti",
              "Yemen", "Sao Tome and Principe", "Sudan", "Eritrea", "Togo",
              "Madagascar", "Haiti", "Eswatini", "Turkmenistan", "Gambia, The",
              "Burundi", "Zambia", "Laos", "Zimbabwe", "Ethiopia",
              "Timor-Leste", "Burma", "Papua New Guinea", "Malawi", "Kiribati",
              "Ghana", "Senegal", "Congo, Republic of the", "India", "Suriname",
              "Tanzania", "Bangladesh", "Uganda", "Cambodia", "Namibia",
              "Tuvalu", "Gabon", "Kenya", "Guatemala", "Rwanda",
              "Kyrgyzstan", "Maldives", "Bhutan", "Nepal", "Botswana",
              "Kosovo", "Cabo Verde", "Bolivia", "Philippines", "Paraguay",
              "South Africa", "Georgia", "Tajikistan", "Dominican Republic", "Guyana",
              "Micronesia, Federated States of", "Marshall Islands", "Mongolia", "Solomon Islands", "Indonesia",
              "Iraq", "Algeria", "Turkey (Turkiye)", "Morocco", "Uzbekistan",
              "Saint Helena, Ascension, and Tristan da Cunha", "Samoa", "Egypt", "Korea, North", "Honduras",
              "Gaza Strip", "Syria", "West Bank", "Trinidad and Tobago", "Cook Islands",
              "Nicaragua", "Iran", "Panama", "Vietnam", "Oman",
              "Venezuela", "Moldova", "Vanuatu", "Antigua and Barbuda", "British Virgin Islands",
              "Jordan", "Brazil", "Saint Vincent and the Grenadines", "Mexico", "Northern Mariana Islands",
              "Tonga", "El Salvador", "Saudi Arabia", "Colombia", "Armenia",
              "Mauritius", "Aruba", "Saint Lucia", "Tunisia", "Belize",
              "Ecuador", "Turks and Caicos Islands", "Azerbaijan", "Guam", "Peru",
              "Palau", "Dominica", "Libya", "Jamaica", "Albania",
              "Seychelles", "Brunei", "Fiji", "Montserrat", "Bahrain",
              "American Samoa", "Barbados", "Bahamas, The", "Grenada", "Argentina",
              "Ukraine", "Greenland", "Croatia", "Cyprus", "Uruguay",
              "Kazakhstan", "Saint Kitts and Nevis", "Saint Pierre and Miquelon", "Bulgaria", "Sint Maarten",
              "Nauru", "Curacao", "Virgin Islands", "Cayman Islands", "Kuwait",
              "North Macedonia", "Sri Lanka", "Costa Rica", "Lebanon", "Saint Martin",
              "Saint Barthelemy", "Russia", "Qatar", "Malaysia", "Thailand",
              "Chile", "China", "San Marino", "Gibraltar", "Puerto Rico",
              "Faroe Islands", "Romania", "United States", "Slovakia", "United Arab Emirates",
              "Bosnia and Herzegovina", "Poland", "New Caledonia", "Latvia", "Hungary",
              "Serbia", "Malta", "Macau", "French Polynesia", "Canada",
              "Isle of Man", "Cuba", "Wallis and Futuna", "Liechtenstein", "United Kingdom",
              "Taiwan", "Jersey", "Lithuania", "Netherlands", "Greece",
              "Estonia", "Guernsey", "Ireland", "New Zealand", "Andorra",
              "Austria", "Montenegro", "Luxembourg", "France", "Belgium",
              "Germany", "Italy", "Anguilla", "Switzerland", "Denmark",
              "Australia", "Korea, South", "Israel", "Czechia", "Hong Kong",
              "Portugal", "Spain", "Sweden", "Belarus", "Bermuda",
              "Finland", "Japan", "Norway", "Monaco", "Iceland",
              "Singapore", "Slovenia"),
  # Column 2: Infant mortality rate (deaths per 1,000 live births, 2024 est.)
  infant_mortality_rate = c(101.3, 83.6, 80.5, 77.4, 71.2,
                            64.3, 62.5, 60.1, 58.2, 57.4,
                            57.4, 55.7, 55.6, 54.9, 53.7,
                            52.9, 52.5, 51.5, 48.9, 47.0,
                            47.0, 46.4, 46.1, 45.7, 45.2,
                            44.6, 42.6, 40.6, 39.8, 38.4,
                            37.5, 36.8, 36.7, 35.9, 35.7,
                            35.7, 35.6, 35.4, 33.4, 32.6,
                            32.2, 32.1, 32.0, 31.9, 31.5,
                            31.2, 31.1, 30.6, 30.4, 29.6,
                            29.6, 28.8, 28.5, 27.9, 27.9,
                            27.8, 26.9, 26.1, 25.0, 24.9,
                            24.5, 24.4, 24.3, 24.0, 23.7,
                            22.9, 22.4, 22.3, 22.0, 22.0,
                            21.9, 21.7, 21.7, 21.7, 21.1,
                            20.9, 20.6, 19.4, 19.1, 18.9,
                            18.7, 18.7, 18.4, 18.3, 18.2,
                            18.1, 17.3, 16.8, 15.4, 15.4,
                            15.1, 15.1, 15.1, 15.1, 15.1,
                            14.4, 14.3, 14.2, 14.1, 13.9,
                            13.9, 13.8, 13.7, 13.6, 13.4,
                            13.2, 12.9, 12.3, 12.1, 11.9,
                            11.8, 11.7, 11.7, 11.7, 11.6,
                            11.6, 11.5, 11.5, 11.3, 11.3,
                            11.2, 11.1, 10.9, 10.9, 10.8,
                            10.8, 10.7, 10.7, 10.7, 10.3,
                            10.2, 10.0, 9.7, 9.7, 9.7,
                            9.7, 9.6, 9.4, 9.0, 9.0,
                            8.7, 8.5, 8.4, 8.1, 8.0,
                            8.0, 8.0, 7.8, 7.7, 7.6,
                            7.6, 7.5, 7.3, 7.3, 7.2,
                            7.0, 6.8, 6.7, 6.7, 6.5,
                            6.5, 6.5, 6.4, 6.4, 6.3,
                            6.3, 6.2, 6.2, 6.0, 5.8,
                            5.7, 5.5, 5.1, 5.1, 5.0,
                            5.0, 4.9, 4.8, 4.7, 4.7,
                            4.5, 4.4, 4.4, 4.3, 4.3,
                            4.1, 4.0, 3.9, 3.9, 3.8,
                            3.8, 3.7, 3.6, 3.6, 3.4,
                            3.3, 3.3, 3.3, 3.3, 3.3,
                            3.2, 3.2, 3.2, 3.1, 3.1,
                            3.1, 3.1, 3.0, 3.0, 3.0,
                            2.9, 2.8, 2.8, 2.6, 2.5,
                            2.4, 2.4, 2.3, 2.1, 2.1,
                            2.1, 1.9, 1.8, 1.7, 1.6,
                            1.5, 1.5),
  # Column 3: Region classification from CIA World Factbook
  region = c("South Asia", "Africa", "Africa", "Africa", "Africa",
             "Africa", "Africa", "Africa", "Africa", "Africa",
             "Africa", "Africa", "Africa", "Africa", "Africa",
             "Africa", "Africa", "South Asia", "Africa", "Africa",
             "Africa", "Africa", "Africa", "Africa", "Africa",
             "Middle East", "Africa", "Africa", "Africa", "Africa",
             "Africa", "Central America and the Caribbean", "Africa", "Central Asia", "Africa",
             "Africa", "Africa", "East and Southeast Asia", "Africa", "Africa",
             "East and Southeast Asia", "East and Southeast Asia", "Australia and Oceania", "Africa", "Australia and Oceania",
             "Africa", "Africa", "Africa", "South Asia", "South America",
             "Africa", "South Asia", "Africa", "East and Southeast Asia", "Africa",
             "Australia and Oceania", "Africa", "Africa", "Central America and the Caribbean", "Africa",
             "Central Asia", "South Asia", "South Asia", "South Asia", "Africa",
             "Europe", "Africa", "South America", "East and Southeast Asia", "South America",
             "Africa", "Middle East", "Central Asia", "Central America and the Caribbean", "South America",
             "Australia and Oceania", "Australia and Oceania", "East and Southeast Asia", "Australia and Oceania", "East and Southeast Asia",
             "Middle East", "Africa", "Middle East", "Africa", "Central Asia",
             "Africa", "Australia and Oceania", "Africa", "East and Southeast Asia", "Central America and the Caribbean",
             "Middle East", "Middle East", "Middle East", "Central America and the Caribbean", "Australia and Oceania",
             "Central America and the Caribbean", "Middle East", "Central America and the Caribbean", "East and Southeast Asia", "Middle East",
             "South America", "Europe", "Australia and Oceania", "Central America and the Caribbean", "Central America and the Caribbean",
             "Middle East", "South America", "Central America and the Caribbean", "North America", "Australia and Oceania",
             "Australia and Oceania", "Central America and the Caribbean", "Middle East", "South America", "Middle East",
             "Africa", "Central America and the Caribbean", "Central America and the Caribbean", "Africa", "Central America and the Caribbean",
             "South America", "Central America and the Caribbean", "Middle East", "Australia and Oceania", "South America",
             "Australia and Oceania", "Central America and the Caribbean", "Africa", "Central America and the Caribbean", "Europe",
             "Africa", "East and Southeast Asia", "Australia and Oceania", "Central America and the Caribbean", "Middle East",
             "Australia and Oceania", "Central America and the Caribbean", "Central America and the Caribbean", "Central America and the Caribbean", "South America",
             "Europe", "North America", "Europe", "Europe", "South America",
             "Central Asia", "Central America and the Caribbean", "North America", "Europe", "Central America and the Caribbean",
             "Australia and Oceania", "Central America and the Caribbean", "Central America and the Caribbean", "Central America and the Caribbean", "Middle East",
             "Europe", "South Asia", "Central America and the Caribbean", "Middle East", "Central America and the Caribbean",
             "Central America and the Caribbean", "Europe", "Middle East", "East and Southeast Asia", "East and Southeast Asia",
             "South America", "East and Southeast Asia", "Europe", "Europe", "Central America and the Caribbean",
             "Europe", "Europe", "North America", "Europe", "Middle East",
             "Europe", "Europe", "Australia and Oceania", "Europe", "Europe",
             "Europe", "Europe", "East and Southeast Asia", "Australia and Oceania", "North America",
             "Europe", "Central America and the Caribbean", "Australia and Oceania", "Europe", "Europe",
             "East and Southeast Asia", "Europe", "Europe", "Europe", "Europe",
             "Europe", "Europe", "Europe", "Australia and Oceania", "Europe",
             "Europe", "Europe", "Europe", "Europe", "Europe",
             "Europe", "Europe", "Central America and the Caribbean", "Europe", "Europe",
             "Australia and Oceania", "East and Southeast Asia", "Middle East", "Europe", "East and Southeast Asia",
             "Europe", "Europe", "Europe", "Europe", "North America",
             "Europe", "East and Southeast Asia", "Europe", "Europe", "Europe",
             "East and Southeast Asia", "Europe")
)
```

```{r}
# Display first 6 rows to verify structure
head(infant_mortality)
```

```{r}
# Verify data types using str() function
str(infant_mortality)
```

```{r}
# Save as RDS format (preserves R data types)
saveRDS(infant_mortality, "infant_mortality.rds")
```

```{r}
# Save as CSV format (plain text, portable)
write.csv(infant_mortality, "infant_mortality.csv", row.names = FALSE)
```

```{r}
# Verify RDS file by reading it back and checking structure
im_rds <- readRDS("infant_mortality.rds")
# Display structure of loaded RDS file
str(im_rds)
```

```{r}
# Verify CSV file by reading it back and checking structure
im_csv <- read.csv("infant_mortality.csv")
# Display structure of loaded CSV file
str(im_csv)
```

## 3. Life Expectancy at Birth Data

```{r}
# Create life expectancy at birth dataframe with 3 columns: country, variable, region
life_expectancy <- data.frame(
  # Column 1: Country names from CIA World Factbook
  country = c("Monaco", "Singapore", "Macau", "Japan", "Canada",
              "San Marino", "Hong Kong", "Iceland", "Switzerland", "Andorra",
              "Guernsey", "Malta", "Australia", "Luxembourg", "Korea, South",
              "Israel", "Jersey", "Spain", "Italy", "Liechtenstein",
              "Sweden", "Norway", "New Zealand", "Austria", "France",
              "Anguilla", "Cayman Islands", "Isle of Man", "Bermuda", "Belgium",
              "Finland", "Slovenia", "United Kingdom", "Puerto Rico", "Denmark",
              "Ireland", "Portugal", "Germany", "Netherlands", "Greece",
              "Saint Pierre and Miquelon", "Faroe Islands", "Taiwan", "Turks and Caicos Islands", "Wallis and Futuna",
              "Saint Martin", "Saint Barthelemy", "Gibraltar", "United States", "Saint Helena, Ascension, and Tristan da Cunha",
              "Virgin Islands", "Bahrain", "Qatar", "Costa Rica", "Chile",
              "Cyprus", "British Virgin Islands", "Cuba", "Curacao", "United Arab Emirates",
              "Albania", "Sint Maarten", "Kuwait", "Saint Lucia", "New Caledonia",
              "Panama", "Lebanon", "Barbados", "French Polynesia", "Uruguay",
              "Brunei", "Argentina", "Paraguay", "Dominica", "China",
              "Czechia", "Bosnia and Herzegovina", "Aruba", "Estonia", "Antigua and Barbuda",
              "Thailand", "Montenegro", "Guam", "Tonga", "Algeria",
              "Croatia", "Libya", "Cook Islands", "Saint Kitts and Nevis", "Maldives",
              "Oman", "North Macedonia", "Tunisia", "Slovakia", "Solomon Islands",
              "Saint Vincent and the Grenadines", "Saudi Arabia", "Northern Mariana Islands", "Romania", "Sri Lanka",
              "Poland", "Armenia", "Turkey (Turkiye)", "Bahamas, The", "Seychelles",
              "Malaysia", "Trinidad and Tobago", "West Bank", "Jordan", "Latvia",
              "Jamaica", "Grenada", "Brazil", "Uzbekistan", "Vietnam",
              "Montserrat", "Lithuania", "Bulgaria", "Hungary", "Azerbaijan",
              "El Salvador", "American Samoa", "Vanuatu", "Samoa", "Iran",
              "Gaza Strip", "Mauritius", "Serbia", "Bangladesh", "Marshall Islands",
              "Palau", "Egypt", "Micronesia, Federated States of", "Colombia", "Ecuador",
              "Fiji", "Syria", "Belarus", "Nicaragua", "Mexico",
              "Greenland", "Venezuela", "Belize", "Cabo Verde", "Morocco",
              "Iraq", "Bhutan", "Indonesia", "Guatemala", "Korea, North",
              "Kazakhstan", "Honduras", "Kosovo", "Nepal", "Malawi",
              "Congo, Republic of the", "Kyrgyzstan", "Georgia", "Suriname", "Dominican Republic",
              "Bolivia", "Turkmenistan", "Guyana", "Russia", "Togo",
              "South Africa", "Tajikistan", "Mongolia", "Cambodia", "Philippines",
              "Tanzania", "Senegal", "Ukraine", "Timor-Leste", "Gabon",
              "Kenya", "Burma", "Pakistan", "Moldova", "Ghana",
              "Papua New Guinea", "Uganda", "Laos", "Tuvalu", "Peru",
              "Madagascar", "Nauru", "Kiribati", "Gambia, The", "India",
              "Yemen", "Burundi", "Comoros", "Sudan", "Sao Tome and Principe",
              "Ethiopia", "Eritrea", "Zimbabwe", "Zambia", "Rwanda",
              "Botswana", "Djibouti", "Namibia", "Mauritania", "Haiti",
              "Guinea", "Guinea-Bissau", "Burkina Faso", "Cameroon", "Equatorial Guinea",
              "Mali", "Cote d'Ivoire", "Benin", "Angola", "Congo, Democratic Republic of the",
              "Nigeria", "Liberia", "Niger", "Eswatini", "South Sudan",
              "Lesotho", "Chad", "Sierra Leone", "Mozambique", "Somalia",
              "Central African Republic", "Afghanistan"),
  # Column 2: Life expectancy at birth (years, 2024 est.)
  life_expectancy_years = c(89.8, 86.7, 85.3, 85.2, 84.2,
                            84.2, 84.0, 84.0, 83.9, 83.8,
                            83.6, 83.6, 83.5, 83.4, 83.4,
                            83.1, 83.0, 83.0, 83.0, 83.0,
                            82.9, 82.9, 82.9, 82.7, 82.6,
                            82.6, 82.5, 82.5, 82.5, 82.3,
                            82.2, 82.2, 82.2, 82.1, 82.1,
                            82.0, 81.9, 81.9, 81.9, 81.9,
                            81.8, 81.7, 81.6, 81.3, 81.1,
                            81.0, 81.0, 80.9, 80.9, 80.9,
                            80.7, 80.4, 80.3, 80.3, 80.3,
                            80.2, 80.1, 80.1, 79.9, 79.9,
                            79.9, 79.7, 79.6, 79.4, 79.3,
                            79.2, 79.2, 79.0, 78.9, 78.9,
                            78.9, 78.8, 78.8, 78.7, 78.7,
                            78.6, 78.5, 78.5, 78.4, 78.3,
                            78.2, 78.2, 78.0, 78.0, 77.9,
                            77.7, 77.7, 77.6, 77.6, 77.4,
                            77.4, 77.3, 77.3, 77.2, 77.2,
                            77.2, 77.2, 77.1, 76.9, 76.8,
                            76.7, 76.7, 76.7, 76.7, 76.6,
                            76.6, 76.5, 76.5, 76.5, 76.4,
                            76.3, 76.3, 76.3, 76.2, 76.1,
                            76.1, 76.1, 76.1, 76.0, 75.9,
                            75.9, 75.8, 75.7, 75.7, 75.6,
                            75.5, 75.4, 75.3, 75.2, 75.2,
                            75.2, 75.0, 75.0, 74.9, 74.9,
                            74.8, 74.8, 74.7, 74.7, 74.6,
                            74.5, 74.5, 74.3, 74.3, 74.2,
                            73.7, 73.7, 73.6, 73.5, 73.5,
                            73.3, 73.1, 73.1, 73.0, 73.0,
                            72.9, 72.9, 72.8, 72.7, 72.6,
                            72.5, 72.4, 72.4, 72.3, 72.1,
                            71.9, 71.9, 71.9, 71.4, 70.8,
                            70.8, 70.6, 70.5, 70.5, 70.4,
                            70.4, 70.3, 70.3, 70.1, 70.1,
                            70.1, 69.7, 69.0, 69.0, 68.9,
                            68.8, 68.6, 68.5, 68.4, 68.2,
                            68.2, 68.1, 67.8, 67.8, 67.7,
                            67.7, 67.5, 67.2, 66.9, 66.6,
                            66.4, 65.9, 65.9, 65.9, 65.6,
                            64.6, 64.5, 64.2, 64.2, 63.9,
                            63.2, 63.2, 63.0, 62.9, 62.6,
                            62.2, 61.6, 60.9, 60.7, 60.3,
                            60.2, 60.0, 59.4, 58.3, 56.5,
                            56.4, 54.4),
  # Column 3: Region classification from CIA World Factbook
  region = c("Europe", "East and Southeast Asia", "East and Southeast Asia", "East and Southeast Asia", "North America",
             "Europe", "East and Southeast Asia", "Europe", "Europe", "Europe",
             "Europe", "Europe", "Australia and Oceania", "Europe", "East and Southeast Asia",
             "Middle East", "Europe", "Europe", "Europe", "Europe",
             "Europe", "Europe", "Australia and Oceania", "Europe", "Europe",
             "Central America and the Caribbean", "Central America and the Caribbean", "Europe", "North America", "Europe",
             "Europe", "Europe", "Europe", "Central America and the Caribbean", "Europe",
             "Europe", "Europe", "Europe", "Europe", "Europe",
             "North America", "Europe", "East and Southeast Asia", "Central America and the Caribbean", "Australia and Oceania",
             "Central America and the Caribbean", "Central America and the Caribbean", "Europe", "North America", "Africa",
             "Central America and the Caribbean", "Middle East", "Middle East", "Central America and the Caribbean", "South America",
             "Europe", "Central America and the Caribbean", "Central America and the Caribbean", "Central America and the Caribbean", "Middle East",
             "Europe", "Central America and the Caribbean", "Middle East", "Central America and the Caribbean", "Australia and Oceania",
             "Central America and the Caribbean", "Middle East", "Central America and the Caribbean", "Australia and Oceania", "South America",
             "East and Southeast Asia", "South America", "South America", "Central America and the Caribbean", "East and Southeast Asia",
             "Europe", "Europe", "Central America and the Caribbean", "Europe", "Central America and the Caribbean",
             "East and Southeast Asia", "Europe", "Australia and Oceania", "Australia and Oceania", "Africa",
             "Europe", "Africa", "Australia and Oceania", "Central America and the Caribbean", "South Asia",
             "Middle East", "Europe", "Africa", "Europe", "Australia and Oceania",
             "Central America and the Caribbean", "Middle East", "Australia and Oceania", "Europe", "South Asia",
             "Europe", "Middle East", "Middle East", "Central America and the Caribbean", "Africa",
             "East and Southeast Asia", "Central America and the Caribbean", "Middle East", "Middle East", "Europe",
             "Central America and the Caribbean", "Central America and the Caribbean", "South America", "Central Asia", "East and Southeast Asia",
             "Central America and the Caribbean", "Europe", "Europe", "Europe", "Middle East",
             "Central America and the Caribbean", "Australia and Oceania", "Australia and Oceania", "Australia and Oceania", "Middle East",
             "Middle East", "Africa", "Europe", "South Asia", "Australia and Oceania",
             "Australia and Oceania", "Africa", "Australia and Oceania", "South America", "South America",
             "Australia and Oceania", "Middle East", "Europe", "Central America and the Caribbean", "North America",
             "North America", "South America", "Central America and the Caribbean", "Africa", "Africa",
             "Middle East", "South Asia", "East and Southeast Asia", "Central America and the Caribbean", "East and Southeast Asia",
             "Central Asia", "Central America and the Caribbean", "Europe", "South Asia", "Africa",
             "Africa", "Central Asia", "Middle East", "South America", "Central America and the Caribbean",
             "South America", "Central Asia", "South America", "Europe", "Africa",
             "Africa", "Central Asia", "East and Southeast Asia", "East and Southeast Asia", "East and Southeast Asia",
             "Africa", "Africa", "Europe", "East and Southeast Asia", "Africa",
             "Africa", "East and Southeast Asia", "South Asia", "Europe", "Africa",
             "Australia and Oceania", "Africa", "East and Southeast Asia", "Australia and Oceania", "South America",
             "Africa", "Australia and Oceania", "Australia and Oceania", "Africa", "South Asia",
             "Middle East", "Africa", "Africa", "Africa", "Africa",
             "Africa", "Africa", "Africa", "Africa", "Africa",
             "Africa", "Africa", "Africa", "Africa", "Central America and the Caribbean",
             "Africa", "Africa", "Africa", "Africa", "Africa",
             "Africa", "Africa", "Africa", "Africa", "Africa",
             "Africa", "Africa", "Africa", "Africa", "Africa",
             "Africa", "Africa", "Africa", "Africa", "Africa",
             "Africa", "South Asia")
)
```

```{r}
# Display first 6 rows to verify structure
head(life_expectancy)
```

```{r}
# Verify data types using str() function
str(life_expectancy)
```

```{r}
# Save as RDS format (preserves R data types)
saveRDS(life_expectancy, "life_expectancy.rds")
```

```{r}
# Save as CSV format (plain text, portable)
write.csv(life_expectancy, "life_expectancy.csv", row.names = FALSE)
```

```{r}
# Verify RDS file by reading it back and checking structure
le_rds <- readRDS("life_expectancy.rds")
# Display structure of loaded RDS file
str(le_rds)
```

```{r}
# Verify CSV file by reading it back and checking structure
le_csv <- read.csv("life_expectancy.csv")
# Display structure of loaded CSV file
str(le_csv)
```

## 4. Summary

### Files Created (6 total):

| Dataset | RDS File | CSV File |
|---------|----------|----------|
| Maternal Mortality Ratio | `maternal_mortality.rds` | `maternal_mortality.csv` |
| Infant Mortality Rate | `infant_mortality.rds` | `infant_mortality.csv` |
| Life Expectancy at Birth | `life_expectancy.rds` | `life_expectancy.csv` |

### Data Sources:
- Maternal Mortality: https://www.cia.gov/the-world-factbook/field/maternal-mortality-ratio/country-comparison/
- Infant Mortality: https://www.cia.gov/the-world-factbook/field/infant-mortality-rate/country-comparison/
- Life Expectancy: https://www.cia.gov/the-world-factbook/field/life-expectancy-at-birth/country-comparison/

```{r}
# List all files created in current directory
list.files(pattern = "\\.(rds|csv)$")
```

